---
title: "MOU_Index"
output: html_document
date: "2024-09-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(gcookbook)
library(ggrepel)
library(ggsci)
library(ggthemes)
library(NLP)
library(ggrepel)
library(worldfootballR)
library(jtools)
library(pwr)
library(ggalt)
```


## Loading Data

```{r data}
DAVIESdata.2024.04.02 <- read.csv("C:/Users/George Ferridge/Downloads/DAVIESdata-2017-2024-raw.csv")
df <- DAVIESdata.2024.04.02
```

## Calculate Team Aggregates on DAVIES data (Minutes weighted)

```{r}
df_group <- df %>% group_by(League, Team, Season, Gen..Role) %>% 
    summarise(DAVIES_TOTAL = sum(DAVIES), 
              DAVIES_weighted = sum(DAVIES*Min)/sum(Min))
```

## Calculate League averages, STDevs and Vars for the same values

```{r}
df_2 <- df_group %>% group_by(League, Season, Gen..Role) %>%
    summarise(League_DAVIES = mean(DAVIES_weighted),
              Raw_League_DAVIES = mean(DAVIES_TOTAL),
              SD_League = sd(DAVIES_weighted),
              Raw_SD_League = sd(DAVIES_TOTAL))
```

## Join these two datasets and create a Z score for each Team and Role

```{r}
df_together <- left_join(df_group,df_2, join_by(League, Season, Gen..Role))


df_together$Z <- (df_together$DAVIES_weighted - df_together$League_DAVIES)/df_together$SD_League
```

 At this point I see that the model is working as intended, with teams coming out that make a lot of sense. Manchester United's weakness in central defense, Barcelona's strength in creativity.


Now I want to add wages to the player data, which I'll do from scraping FBRef. This code does  a few different things. It firstly takes the nations that I would like it to (top 5 leagues) and gets league URLS for it for every year in the sample. It then scrapes these league sites for the team urls. It then scrapes wage data for all available team_urls, converts it into a single value from its initial state as a length one vector and then saves it as a csv which will be called instead from now on. This code can be skipped if there is a link to the csv. 

```{r, eval=FALSE, echo=FALSE}


nations <- c("GER", "ENG", "FRA", "ESP", "ITA")

league_urls <- c()

for (year in 2018:2024){
  for (nation in nations){
    new_urls <- fb_league_urls(country=nation, gender = "M", year, tier = "1st")
    league_urls <- c(league_urls, new_urls)
  }
}

team_urls <- c()
for (league in league_urls){
  new_urls <- fb_teams_urls(league, time_pause=3)
  team_urls <- c(team_urls, new_urls)
}


team_urls[team_urls != "https://fbref.com/en/squads/60b5e41f/2017-2018/Hannover-96-Stats"]

wage_data = data.frame()

get_wages <- function(urls, df){
  tryCatch(
    {
      for (url in urls){
        stringurl <- as.String(url)
        new_df <- fb_squad_wages(team_urls = url, time_pause = 5)
        df <- rbind(df, new_df)
      }
      error = function(e) {
        print(stringurl + " failed")
        print(e)
      }
      return(df)
    }
  )
}
wage_data <- get_wages(team_urls, wage_data)


wage_data = data.frame()

get_wages <- function(urls, df){
  results_list <- list()  # Use a list to store data frames
  for (url in urls) {
    tryCatch({
      new_df <- fb_squad_wages(team_urls = url, time_pause = 5)
      results_list[[url]] <- new_df
    }, error = function(e) {
      print(paste(url, "failed"))
      print(e$message)  # Print error message
    })
  }
  # Combine all data frames into one
  df <- do.call(rbind, results_list)
  return(df)
}

# Assuming team_urls is defined somewhere in your script
wage_data <- get_wages(team_urls, wage_data)

wage_data$WeeklyWageEUR <- as.String(wage_data$WeeklyWageEUR)

wage_data$WeeklyWageEUR <- sapply(wage_data$WeeklyWageEUR, function(x) { x[[1]] })
wage_data$WeeklyWageGBP <- sapply(wage_data$WeeklyWageGBP, function(x) { x[[1]] })
wage_data$WeeklyWageUSD <- sapply(wage_data$WeeklyWageUSD, function(x) { x[[1]] })
wage_data$AnnualWageEUR <- sapply(wage_data$AnnualWageEUR, function(x) { x[[1]] })
wage_data$AnnualWageGBP <- sapply(wage_data$AnnualWageGBP, function(x) { x[[1]] })
wage_data$AnnualWageUSD <- sapply(wage_data$AnnualWageUSD, function(x) { x[[1]] })

write.csv(wage_data, "wage_data.csv", row.names=FALSE)

```


## Use preprepared wage data instead of scraping

```{r}
wage_data <- read.csv("wage_data.csv")
```


## Transformations to alter wage data storage for future manipulation

```{r}

wage_data[] <- lapply(wage_data, function(x) if (is.list(x)) sapply(x, function(y) y[[1]], simplify = "vector") else x)

sapply(wage_data, is.list)

wage_data$WeeklyWageEUR <- sapply(wage_data$WeeklyWageEUR, function(x) {
  if (length(x) == 0 || is.null(x[[1]])) NA else as.numeric(x[[1]])
})

wage_data$WeeklyWageEUR <- vapply(wage_data$WeeklyWageEUR, function(x) {
  if (length(x) > 0 && !is.null(x[[1]])) as.numeric(x[[1]]) else NA
}, numeric(1))

wage_data$WeeklyWageGBP<- sapply(wage_data$WeeklyWageGBP, function(x) {
  if (length(x) == 0 || is.null(x[[1]])) NA else as.numeric(x[[1]])
})

wage_data$WeeklyWageGBP <- vapply(wage_data$WeeklyWageGBP, function(x) {
  if (length(x) > 0 && !is.null(x[[1]])) as.numeric(x[[1]]) else NA
}, numeric(1))

wage_data$WeeklyWageUSD <- sapply(wage_data$WeeklyWageUSD, function(x) {
  if (length(x) == 0 || is.null(x[[1]])) NA else as.numeric(x[[1]])
})

wage_data$WeeklyWageUSD <- vapply(wage_data$WeeklyWageUSD, function(x) {
  if (length(x) > 0 && !is.null(x[[1]])) as.numeric(x[[1]]) else NA
}, numeric(1))

wage_data$AnnualWageEUR <- sapply(wage_data$AnnualWageEUR, function(x) {
  if (length(x) == 0 || is.null(x[[1]])) NA else as.numeric(x[[1]])
})

wage_data$AnnualWageEUR <- vapply(wage_data$AnnualWageEUR, function(x) {
  if (length(x) > 0 && !is.null(x[[1]])) as.numeric(x[[1]]) else NA
}, numeric(1))

wage_data$AnnualWageGBP <- sapply(wage_data$AnnualWageGBP, function(x) {
  if (length(x) == 0 || is.null(x[[1]])) NA else as.numeric(x[[1]])
})

wage_data$AnnualWageGBP <- vapply(wage_data$AnnualWageGBP, function(x) {
  if (length(x) > 0 && !is.null(x[[1]])) as.numeric(x[[1]]) else NA
}, numeric(1))

wage_data$AnnualWageUSD <- sapply(wage_data$AnnualWageUSD, function(x) {
  if (length(x) == 0 || is.null(x[[1]])) NA else as.numeric(x[[1]])
})

wage_data$AnnualWageUSD <- vapply(wage_data$AnnualWageUSD, function(x) {
  if (length(x) > 0 && !is.null(x[[1]])) as.numeric(x[[1]]) else NA
}, numeric(1))

wage_data$WeeklyWageEUR <- wage_data$WeeklyWageEUR_num
```

## Add wage data to the original dataframe by player and season

```{r}
df_wage <- left_join(df, wage_data, by=c("Player", "Season"))
```


## Add in lags variables for previous seasons performance in order to calibrate models
```{r}
lags <- df_wage

lags <- lags %>%
  arrange(Player, Season) %>%  # Sort by player and season
  group_by(Player) %>%         # Group by player, but not reducing data
  mutate(
    stats_last_season = lag(DAVIES, 1),     # Stats from the previous season
    stats_2_seasons_ago = lag(DAVIES, 2),   # Stats from two seasons ago
    stats_3_seasons_ago = lag(DAVIES, 3),   # Stats from three seasons ago
    stats_4_seasons_ago = lag(DAVIES, 4),   # Stats from four seasons ago
    stats_5_seasons_ago = lag(DAVIES, 5),   # Stats from five seasons ago
    stats_6_seasons_ago = lag(DAVIES, 6)    # Stats from six seasons ago
  ) %>%
  ungroup()  # Optionally remove the grouping structure
```


## A couple of prediction models, selected from a larger suite. Linear Model 2 is used as prediction algorithm moving forward

```{r}
lm1 <- lm(DAVIES ~ WeeklyWageEUR+League+Gen..Role, df_wage)

summary(lm1)

lm2 <- lm(DAVIES ~ WeeklyWageEUR+Gen..Role+League+Min, lags)

summary(lm2)
```

## Generate Predictions for as many players as possible. Debatable assumption made here: players who do not have data from the previous season are assumed to have been purely average for their age, scoring a 0.

```{r}

lags2 <- lags

interested_leagues <- c("Premier League", "La Liga", "Serie A", "Bundesliga", "Ligue 1")

# Filter the dataframe
filtered_lags2 <- lags2 %>%
  filter(League %in% interested_leagues)

# Display the filtered dataframe
df_predict <- filtered_lags2 %>%
  mutate(across(starts_with("stats"), ~replace_na(., 0)))

# Predict using the model
df_predict$predicted_stats <- predict(lm2, newdata = filtered_lags2, type = "response")

```


## Grouping these predictions by position group and league averages, much like I did for current season performance.

```{r}
grouped_predictions <- df_predict %>% group_by(League, Team.x, Season, Gen..Role) %>% 
  summarise(predicted_stats = sum(predicted_stats, na.rm=TRUE), 
            weighted_predictions = sum(predicted_stats*Min, na.rm=TRUE)/sum(Min, na.rm=TRUE))

league_predictions <- grouped_predictions %>% group_by(League, Season, Gen..Role) %>%
  summarise(League_predictions = mean(weighted_predictions),
            Raw_League_predictions = mean(predicted_stats),
            SD_predictions = sd(weighted_predictions),
            Raw_SD_predictions = sd(predicted_stats))


predictions_together <- left_join(grouped_predictions, league_predictions, join_by(League, Season, Gen..Role))
```

## Slight cleaning to make future analysis easier

```{r}

predictions_together$Z_predicted <- ifelse(predictions_together$SD_predictions > 0, (predictions_together$weighted_predictions - predictions_together$League_predictions)/predictions_together$SD_predictions, 0)

predictions_together$Team <- predictions_together$Team.x
```

## Combined data

```{r}
data_plus_predictions <- left_join(df_together, predictions_together, join_by(Team, Season, Gen..Role))
```


## Importing manager data

```{r}
Managers_DAVIES <- read.csv("Managers_DAVIES.csv")
data_plus_predictions$League <- data_plus_predictions$League.x
managers <- left_join(Managers_DAVIES, data_plus_predictions, join_by(League, Gen..Role, Team, Season))
```


## Linear models to look at the effect of individual managers

```{r}

lm5 <- lm(Z.x ~ Manager+Z_predicted, data=managers)
lm6 <- lm(DAVIES_weighted.y ~ Manager + weighted_predictions, managers)

summary(lm5)

lm6 <- lm(DAVIES_weighted.y ~ weighted_predictions, managers)

plot_summs(lm6, coefs = c( "ManagerPep Guardiola", "ManagerJurgen Klopp", "ManagerMichel","ManagerRoberto De Zerbi",
                           "ManagerMauricio Pochettino", "ManagerXavi", "ManagerXabi Alonso", "ManagerSteve Bruce",
                           "ManagerScott Parker", "ManagerRoy Hodgson", "ManagerMikel Arteta", "ManagerJose Mourinho",
                           "ManagerErik ten Hag", "ManagerChris Wilder", "ManagerChris Hughton", "ManagerGraham Potter"))
summary(lm6)

ggsave(filename="SelectedManagers.png", width=11, height=7, dpi=1200)


```



